---
title: "Assignment 1"
author: "Spandana Sodadasi"
date: "2023-09-06"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE, comment = NA)
```

**LP Model**

1. Back Savers is a company that produces backpacks primarily for students. They are 
considering offering some combination of two different models—the Collegiate and the 
Mini. Both are made out of the same rip-resistant nylon fabric. Back Savers has a long term contract with a supplier of the nylon and receives a 5000 square-foot shipment of 
the material each week. Each Collegiate requires 3 square feet while each Mini requires 2 
square feet. The sales forecasts indicate that at most 1000 Collegiates and 1200 Minis can 
be sold per week. Each Collegiate requires 45 minutes of labor to produce and generates 
a unit profit of $32. Each Mini requires 40 minutes of labor and generates a unit profit of
$24. Back Savers has 35 laborers that each provides 40 hours of labor per week. 
Management wishes to know what quantity of each type of backpack to produce per 
week.

```{r}
data= matrix(c(2,1000,'0.75',3,1200,'0.67'), ncol=3, byrow=TRUE)

# specify the column names and row names of matrix

colnames(data) = c('Material','Sales','Labor')
rownames(data) <- c('Collegiate','Mini')

# assign to table
final=as.table(data)
 
# display
final

```
The management wants to know the quantity of each type of backpack that is to be produced per week in order to maximize their profit.

Assume,

The number of collegiate backpacks produced per week $$ = B_c$$
The number of mini backpacks produced per week $$ = B_m$$ 
(a) The Decision Variables are $$ = B_c \hspace{.2cm} and \hspace{.2cm} B_m$$
(b) Objective function is to Maximize the Profit.

 $$Max \hspace{.3cm} Z = 32B_c+ 24B_m $$
(c) Constraints:- 


Non-negativity of the Decision Variables: $$ B_c \ge 0 \hspace{.2cm} and \hspace{.2cm} B_m \ge 0 $$ 
(d) The full Mathematical formulation for the LP problem is as follows,

$$Max \hspace{.3cm} Z = 32B_c+ 24B_m $$Material constraint: $$ 3B_c + 2B_m \le 5000   $$
Sales constraint: $$ B_c \le 1000 $$ $$ B_m \le 1200 $$
Labor (Hours) Constraint: $$ 0.75B_c + 0.67B_m \le 1400 $$
Such that,

Material Constraint: $$ 3B_c + 2B_m \le 5000   $$
Sales Constraint: $$ B_c \le 1000 $$ $$ B_m \le 1200 $$
Labor (Hours) Constraint: $$ 0.75B_c + 0.67B_m \le 1400 $$
Non-negativity of the Decision Variables: $$ B_c \ge 0 \hspace{.2cm} and \hspace{.2cm} B_m \ge 0 $$ 



2. The Weigelt Corporation has three branch plants with excess production capacity. 
Fortunately, the corporation has a new product ready to begin production, and all three 
plants have this capability, so some of the excess capacity can be used in this way. This 
product can be made in three sizes--large, medium, and small--that yield a net unit profit 
of $420, $360, and $300, respectively. Plants 1, 2, and 3 have the excess capacity to 
produce 750, 900, and 450 units per day of this product, respectively, regardless of the 
size or combination of sizes involved.
The amount of available in-process storage space also imposes a limitation on the 
production rates of the new product. Plants 1, 2, and 3 have 13,000, 12,000, and 5,000 
square feet, respectively, of in-process storage space available for a day's production of 
this product. Each unit of the large, medium, and small sizes produced per day requires 
20, 15, and 12 square feet, respectively.
Sales forecasts indicate that if available, 900, 1,200, and 750 units of the large, 
medium, and small sizes, respectively, would be sold per day.
At each plant, some employees will need to be laid off unless most of the plant’s 
excess production capacity can be used to produce the new product. To avoid layoffs if 
possible, management has decided that the plants should use the same percentage of their 
excess capacity to produce the new product.
Management wishes to know how much of each of the sizes should be produced 
by each of the plants to maximize profit.

```{r}
data= matrix(c(1,750,900,'$420',2,900,1200,'$360',3,450,750,'$300'), ncol=4, byrow=TRUE)

# specify the column names and row names of matrix

colnames(data) = c('Plant','Excess Capacity','Sales Forecast','Profit per unit')
rownames(data) <- c('Large','Medium','Small')

# assign to table
final=as.table(data)
 
# display
final
  
```

```{r}
data= matrix(c(1,20,15,12,'13000',2,20,15,12,'12000',3,20,15,12,'5000'), ncol=5, byrow=TRUE)

# specify the column names and row names of matrix

colnames(data) = c('Plant','Large','Medium','Small','Available Storage Space')
rownames(data) <- NULL

# assign to table
final=as.table(data)
 
# display
final

```
  

The Management wishes to know how much of each of the sizes should be produced 
by each of the plants to maximize profit.

Assume,

The number of large size units produces per day at Plant 1 $$ = P1_L$$
The number of medium size units produces per day at Plant 1 $$ = P1_M$$
The number of small size units produces per day at Plant 1 $$ = P1_S$$
The number of large size units produces per day at Plant 2 $$ = P2_L$$
The number of medium size units produces per day at Plant 2 $$ = P2_M$$
The number of small size units produces per day at Plant 2 $$ = P2_S$$
The number of large size units produces per day at Plant 3 $$ = P3_L$$
The number of medium size units produces per day at Plant 3 $$ = P3_M$$
The number of small size units produces per day at Plant 3 $$ = P3_S$$

(a) So the Decision Variables are $$ = P1_L,P1_M,P1_S,P2_L,P2_M,P2_S,P3_L,P3_M \hspace{.2cm} and \hspace{.2cm} P3_S$$

(b) Objective function is to maximize the profit.

 $$Max \hspace{.3cm} Z = 420P1_L+360P1_M+300P1_S+420P2_L+360P2_M+300P2_S+420P3_L+360P3_M+300P3_S $$
 
(c) Constraints:- 

Capacity constraint: $$ P1_L+P1_M+P1_S \le 750   $$
                     $$ P2_L+P2_M+P2_S \le 900   $$
                     $$ P3_L+P3_M+P3_S \le 450   $$

Storage Space Constraint: $$ 20P1_L+15P1_M+12P1_S \le 13000   $$
                          $$ 20P2_L+15P2_M+12P2_S \le 12000   $$
                          $$ 20P3_L+15P3_M+12P3_S \le 5000   $$

Sales Constraint: $$ P1_L+P2_L+P3_L \le 900   $$
                  $$ P1_M+P2_M+P3_M \le 1200   $$
                  $$ P1_S+P2_S+P3_S \le 750   $$
            
Excess Capacity constraint:
  In this constraint, we need to check whether the same percentage of excess capacity is used to produce the new product at all three plants.
                  $$ P1_L+P1_M+P1_S/750 \hspace{.3cm} - \hspace{.3cm} P2_L+P2_M+P2_S/900 \hspace{.3cm} = \hspace{.3cm} 0  $$
                  $$ P2_L+P2_M+P2_S/900 \hspace{.3cm} - \hspace{.3cm} P3_L+P3_M+P3_S/450 \hspace{.3cm} = \hspace{.3cm} 0  $$
                  $$ P1_L+P1_M+P1_S/750 \hspace{.3cm} - \hspace{.3cm} P3_L+P3_M+P3_S/450 \hspace{.3cm} = \hspace{.3cm} 0  $$

Non-negativity of the Decision Variables: $$P1_L\ge 0, P1_M\ge 0, P1_S\ge 0, P2_L\ge 0, P2_M\ge 0, P2_S\ge 0, P3_L\ge 0, P3_M\ge 0, P3_S\ge 0   $$

(d) The full LP model of the given problem is as follows,

$$Max \hspace{.3cm} Z = 420P1_L+360P1_M+300P1_S+420P2_L+360P2_M+300P2_S+420P3_L+360P3_M+300P3_S $$

Such that,

Capacity constraint: $$ P1_L+P1_M+P1_S \le 750   $$
                     $$ P2_L+P2_M+P2_S \le 900   $$
                     $$ P3_L+P3_M+P3_S \le 450   $$

Storage Space Constraint: $$ 20P1_L+15P1_M+12P1_S \le 13000   $$
                          $$ 20P2_L+15P2_M+12P2_S \le 12000   $$
                          $$ 20P3_L+15P3_M+12P3_S \le 5000   $$

Sales Constraint: $$ P1_L+P2_L+P3_L \le 900   $$
                  $$ P1_M+P2_M+P3_M \le 1200   $$
                  $$ P1_S+P2_S+P3_S \le 750   $$
            
Excess Capacity constraint:
  In this constraint, we need to check whether the same percentage of excess capacity is used to produce the new product at all three plants.
                  $$ P1_L+P1_M+P1_S/750 \hspace{.3cm} - \hspace{.3cm} P2_L+P2_M+P2_S/900 \hspace{.3cm} = \hspace{.3cm} 0  $$
                  $$ P2_L+P2_M+P2_S/900 \hspace{.3cm} - \hspace{.3cm} P3_L+P3_M+P3_S/450 \hspace{.3cm} = \hspace{.3cm} 0  $$
                  $$ P1_L+P1_M+P1_S/750 \hspace{.3cm} - \hspace{.3cm} P3_L+P3_M+P3_S/450 \hspace{.3cm} = \hspace{.3cm} 0  $$

Non-negativity of the Decision Variables: $$P1_L\ge 0, P1_M\ge 0, P1_S\ge 0, P2_L\ge 0, P2_M\ge 0, P2_S\ge 0, P3_L\ge 0, P3_M\ge 0, P3_S\ge 0   $$


