---
title: "QMM Assignment - 2"
author: "Spandana Sodadasi"
date: "2023-09-20"
output: html_document
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE, comment = NULL)
```

The Weigelt Corporation has three branch plants with excess production capacity. 
Fortunately, the corporation has a new product ready to begin production, and all three 
plants have this capability, so some of the excess capacity can be used in this way. This 
product can be made in three sizes--large, medium, and small--that yield a net unit profit 
of $420, $360, and $300, respectively. Plants 1, 2, and 3 have the excess capacity to 
produce 750, 900, and 450 units per day of this product, respectively, regardless of the 
size or combination of sizes involved.
The amount of available in-process storage space also imposes a limitation on the 
production rates of the new product. Plants 1, 2, and 3 have 13,000, 12,000, and 5,000 
square feet, respectively, of in-process storage space available for a day's production of 
this product. Each unit of the large, medium, and small sizes produced per day requires 
20, 15, and 12 square feet, respectively.
Sales forecasts indicate that if available, 900, 1,200, and 750 units of the large, 
medium, and small sizes, respectively, would be sold per day.
At each plant, some employees will need to be laid off unless most of the plant’s 
excess production capacity can be used to produce the new product. To avoid layoffs if 
possible, management has decided that the plants should use the same percentage of their 
excess capacity to produce the new product.
Management wishes to know how much of each of the sizes should be produced 
by each of the plants to maximize profit.

```{r}
data= matrix(c(1,750,900,'$420',2,900,1200,'$360',3,450,750,'$300'), ncol=4, byrow=TRUE)

# specify the column names and row names of matrix

colnames(data) = c('Plant','Excess Capacity','Sales Forecast','Profit per unit')
rownames(data) <- c('Large','Medium','Small')

# assign to table
final=as.table(data)
 
# display
final
  
```

```{r}
data= matrix(c(1,20,15,12,'13000',2,20,15,12,'12000',3,20,15,12,'5000'), ncol=5, byrow=TRUE)

# specify the column names and row names of matrix

colnames(data) = c('Plant','Large','Medium','Small','Available Storage Space')
rownames(data) <- NULL

# assign to table
final=as.table(data)
 
# display
final

```
  

The Management wishes to know how much of each of the sizes should be produced 
by each of the plants to maximize profit.

Assume,

The number of large size units produces per day at Plant 1 $$ = P1_L$$
The number of medium size units produces per day at Plant 1 $$ = P1_M$$
The number of small size units produces per day at Plant 1 $$ = P1_S$$
The number of large size units produces per day at Plant 2 $$ = P2_L$$
The number of medium size units produces per day at Plant 2 $$ = P2_M$$
The number of small size units produces per day at Plant 2 $$ = P2_S$$
The number of large size units produces per day at Plant 3 $$ = P3_L$$
The number of medium size units produces per day at Plant 3 $$ = P3_M$$
The number of small size units produces per day at Plant 3 $$ = P3_S$$

(a) So the Decision Variables are $$ = P1_L,P1_M,P1_S,P2_L,P2_M,P2_S,P3_L,P3_M \hspace{.2cm} and \hspace{.2cm} P3_S$$

(b) Objective function is to maximize the profit.

 $$Max \hspace{.3cm} Z = 420P1_L+360P1_M+300P1_S+420P2_L+360P2_M+300P2_S+420P3_L+360P3_M+300P3_S $$
 
(c) Constraints:- 

Capacity constraint: $$ P1_L+P1_M+P1_S \le 750   $$
                     $$ P2_L+P2_M+P2_S \le 900   $$
                     $$ P3_L+P3_M+P3_S \le 450   $$

Storage Space Constraint: $$ 20P1_L+15P1_M+12P1_S \le 13000   $$
                          $$ 20P2_L+15P2_M+12P2_S \le 12000   $$
                          $$ 20P3_L+15P3_M+12P3_S \le 5000   $$

Sales Constraint: $$ P1_L+P2_L+P3_L \le 900   $$
                  $$ P1_M+P2_M+P3_M \le 1200   $$
                  $$ P1_S+P2_S+P3_S \le 750   $$
            
Excess Capacity constraint:
  In this constraint, we need to check whether the same percentage of excess capacity is used to produce the new product at all three plants.
                  $$ P1_L+P1_M+P1_S/750 \hspace{.3cm} - \hspace{.3cm} P2_L+P2_M+P2_S/900 \hspace{.3cm} = \hspace{.3cm} 0  $$
                  $$ P2_L+P2_M+P2_S/900 \hspace{.3cm} - \hspace{.3cm} P3_L+P3_M+P3_S/450 \hspace{.3cm} = \hspace{.3cm} 0  $$
                  $$ P1_L+P1_M+P1_S/750 \hspace{.3cm} - \hspace{.3cm} P3_L+P3_M+P3_S/450 \hspace{.3cm} = \hspace{.3cm} 0  $$

Non-negativity of the Decision Variables: $$P1_L\ge 0, P1_M\ge 0, P1_S\ge 0, P2_L\ge 0, P2_M\ge 0, P2_S\ge 0, P3_L\ge 0, P3_M\ge 0, P3_S\ge 0   $$

(d) The full LP model of the given problem is as follows,

$$Max \hspace{.3cm} Z = 420P1_L+360P1_M+300P1_S+420P2_L+360P2_M+300P2_S+420P3_L+360P3_M+300P3_S $$

Such that,

Capacity constraint: $$ P1_L+P1_M+P1_S \le 750   $$
                     $$ P2_L+P2_M+P2_S \le 900   $$
                     $$ P3_L+P3_M+P3_S \le 450   $$

Storage Space Constraint: $$ 20P1_L+15P1_M+12P1_S \le 13000   $$
                          $$ 20P2_L+15P2_M+12P2_S \le 12000   $$
                          $$ 20P3_L+15P3_M+12P3_S \le 5000   $$

Sales Constraint: $$ P1_L+P2_L+P3_L \le 900   $$
                  $$ P1_M+P2_M+P3_M \le 1200   $$
                  $$ P1_S+P2_S+P3_S \le 750   $$
            
Excess Capacity constraint:
  In this constraint, we need to check whether the same percentage of excess capacity is used to produce the new product at all three plants.
                  $$ P1_L+P1_M+P1_S/750 \hspace{.3cm} - \hspace{.3cm} P2_L+P2_M+P2_S/900 \hspace{.3cm} = \hspace{.3cm} 0  $$
                  $$ P2_L+P2_M+P2_S/900 \hspace{.3cm} - \hspace{.3cm} P3_L+P3_M+P3_S/450 \hspace{.3cm} = \hspace{.3cm} 0  $$
                  $$ P1_L+P1_M+P1_S/750 \hspace{.3cm} - \hspace{.3cm} P3_L+P3_M+P3_S/450 \hspace{.3cm} = \hspace{.3cm} 0  $$

Non-negativity of the Decision Variables: $$P1_L\ge 0, P1_M\ge 0, P1_S\ge 0, P2_L\ge 0, P2_M\ge 0, P2_S\ge 0, P3_L\ge 0, P3_M\ge 0, P3_S\ge 0   $$

## Solving the problem using lpSolve Package.

#Installing the lpSolve package.
```{r}
#install.packages("lpSolve")
library(lpSolve)

```

#Setting an Objective function.
```{r}
f.obj<-c(420,360,300,420,360,300,420,360,300)
```

#Setting the Constraint Matrix.
```{r}
f.con<- matrix(c(1,1,1,0,0,0,0,0,0,
                0,0,0,1,1,1,0,0,0,
                0,0,0,0,0,0,1,1,1,
                20,15,12,0,0,0,0,0,0,
                0,0,0,20,15,12,0,0,0,
                0,0,0,0,0,0,20,15,12,
                1,0,0,1,0,0,1,0,0,
                0,1,0,0,1,0,0,1,0,
                0,0,1,0,0,1,0,0,1,
                900,900,900,-750,-750,-750,0,0,0,
                0,0,0,450,450,450,-900,-900,-900,
                450,450,450,0,0,0,-750,-750,-750),nrow=12,byrow= TRUE)
```

#Setting the Inequality Signs.
```{r}
f.dir<-c("<=",
         "<=",
         "<=",
         "<=",
         "<=",
         "<=",
         "<=",
         "<=",
         "<=",
         "=",
         "=",
         "=")
```

#Setting the right hand side coefficients.
```{r}
f.rhs<-c(750,900,450,13000,12000,5000,900,1200,750,0,0,0)
```

#Getting the value of Objection function.
```{r}
opt=lp("max", f.obj,f.con,f.dir,f.rhs)
opt$objval
```

#Getting the values of the Decision Variables.
```{r}
opt$solution[1]
opt$solution[2]
opt$solution[3]
opt$solution[4]
opt$solution[5]
opt$solution[6]
opt$solution[7]
opt$solution[8]
opt$solution[9]

```