---
title: "Assignment: Final Examination"
author: "Group 10"
date: "2023-11-23"
output: html_document
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE, comment = NULL)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
```

#### *<b><u>Problem Statement:-</b></u>*

*You have been assigned the task of forming groups. Consider that your class consists of 15 students, and you aim to create 5 groups, each containing 3 students. The main goal is to maximize the chance of success for each group on a class project. The requirements for forming groups are as follows:*

*• Divide 15 students into 5 groups of 3 members each.*                                                             
*• What factors affect the success of groups? Define three factors (e.g., GPA, gender, etc...)that you feel influence the contributions students make towards project success.*  

*• Use R's randomization function to generate data for these factors.*                                             
*• Record the collected data.*                                                                                     
*• Formulate a mathematical model with an objective function, constraints, and non-negativity components.*         
*• Solve the model in R and report objective function values and decision variables.*                               
*• Provide insights in your results.* 

#### *<b><u>Solution:-</b></u>*

*In order to form five groups for a group project, the objective is to strategically distribute 15 students based on their individual factors.*

*The following are the factors of the students that play a role in dividing them into distinct groups:*

*<b><u>1. GPA of the student:</b></u> It reflects the average performance of a student in their academic courses. It serves as a concise measure of a student's overall success in their studies and provides insights into their academic standing. It plays a key role in the project's success by showcasing the student's quality and capabilities.*

*<b><u>2. Technical Skills:</b></u> Technical skills indicate a student's proficiency in specific technical areas relevant to the project. These skills reflect their ability to contribute effectively and perform tasks that require technical expertise.*

*<b><u>3. Teamwork:</b></u> Teamwork reflects a student's ability to collaborate with others, share responsibilities, and contribute positively to group dynamics. It indicates how well a student can work collectively with peers to achieve common goals, fostering a cooperative and productive group environment.*

*Proportion (out of 100) of the factors contributed for the group project success:*

*<b>1. GPA of the student:</b> 40%*

*<b>2. Technical Skills:</b> 30%*

*<b>3. Teamwork:</b> 30%*

##### *Using R's randomization function to generate data for these factors.*
```{r}
set.seed(123)
num_rows <- 15

GPA <- round(runif(num_rows, min = 2.5, max = 4.0),2)
gpa_data <- data.frame(GPA = GPA)

Technical_skill_scale <- round(runif(num_rows, min = 4, max = 10),0)
gpa_data <- data.frame(Technical_skill_scale = Technical_skill_scale)

Teamwork_scale <- round(runif(num_rows, min = 4, max = 10),0)
gpa_data <- data.frame(Teamwork_scale = Teamwork_scale)

GPA_mean <- mean(GPA)
round(GPA_mean,2)
Technical_skill_mean <- mean(Technical_skill_scale)
round(Technical_skill_mean,0)
Teamwork_mean <- mean(Teamwork_scale)
round(Teamwork_mean,0)

library(kableExtra, warn.conflicts = FALSE)
Student <- c("1","2","3","4","5","6","7","8","9","10","11","12","13","14","15","Avg")
GPA_Value <- c(GPA, 3.35)
Technical_skill_scale_Value <- c(Technical_skill_scale, 7)
Teamwork_scale_value <- c(Teamwork_scale, 7)

data_1 <- data.frame(Student,GPA_Value, Technical_skill_scale_Value, Teamwork_scale_value)

gpa_weight <- 0.4
technical_skill_weight <- 0.3
teamwork_weight <- 0.3

data_1$Weighted_Avg <- round((data_1$GPA_Value * gpa_weight +
                        data_1$Technical_skill_scale_Value * technical_skill_weight +
                        data_1$Teamwork_scale_value * teamwork_weight),2)

data_1 %>%
  kable(align = "c", format = "html") %>%
  kable_styling(bootstrap_options = "bordered", full_width = TRUE) %>%
  row_spec(16,bold = TRUE)%>%
    row_spec(15,extra_css = "border-bottom:dotted;")%>%
add_header_above(c("Factors Contributing for Success of the Group" = 5))
```

#### *<b>1.Formulate the Model: Define the decision variables, objective function, and constraints based on the given requirements and preliminary issues.</b>*

*Before getting into the formulation we first need to define the decision variables which is discussed as follows,*   

$$ G_i \hspace{0.2cm} represent  \hspace{0.2cm} groups \hspace{0.2cm} where \hspace{0.2cm} i=1,2,3,4,5 $$

$$ S_j \hspace{0.2cm} represents \hspace{0.2cm} students \hspace{0.2cm} where \hspace{0.2cm} j=1,2,3,4,5,6,7,8,9,10,11,12,13,14,15 $$

#### *<b><u>Objective function:-</b></u>*

*The primary goal is to optimize group success in a class project, taking into account three factors—GPA, Technical Skill, and Teamwork. These factors, identified as influential in student contributions to project success, have assigned weights for their respective contributions where GPA holds 40%, Technical Skill 30%, and Teamwork 30%. In formulating the objective function, we computed the weighted average of these three factors. The resulting objective function is as follows,*

$$ Max \hspace{0.2cm} Z = 6.87G_1S_1 +  5.67G_1S_2 + 4.84G_1S_3 + 6.03G_1S_4 + 5.76G_1S_5 + 5.83G_1S_6 + 6.42G_1S_7 + 5.44G_1S_8 + 6.13G_1S_9 + 5.17G_1S_{10} + 5.48G_1S_{11} + 5.17G_1S_{12} + 5.61G_1S_{13} + 4.94G_1S_{14} + 4.06G_1S_{15} $$ 
$$ + 6.87G_2S_1 + 5.67G_2S_2 + 4.84G_2S_3 + 6.03G_2S_4 + 5.76G_2S_5 + 5.83G_2S_6 + 6.42G_2S_7 + 5.44G_2S_8 + 6.13G_2S_9 + 5.17G_2S_{10} + 5.48G_2S_{11} + 5.17G_2S_{12} + 5.61G_2S_{13} + 4.94G_2S_{14} + 4.06G_2S_{15} $$ 
$$ + 6.87G_3S_1 + 5.67G_3S_2 + 4.84G_3S_3 + 6.03G_3S_4 + 5.76G_3S_5 + 5.83G_3S_6 + 6.42G_3S_7 + 5.44G_3S_8 + 6.13G_3S_9 + 5.17G_3S_{10} + 5.48G_3S_{11} +  5.17G_3S_{12} + 5.61G_3S_{13} + 4.94G_3S_{14} + 4.06G_3S_{15} $$ 
$$ + 6.87G_4S_1 + 5.67G_4S_2 + 4.84G_4S_3 + 6.03G_4S_4 + 5.76G_4S_5 + 5.83G_4S_6 + 6.42G_4S_7 + 5.44G_4S_8 + 6.13G_4S_9 + 5.17G_4S_{10} + 5.48G_4S_{11} + 5.17G_4S_{12} + 5.61G_4S_{13} + 4.94G_4S_{14} + 4.06G_4S_{15} $$ 
$$ + 6.87G_5S_1 + 5.67G_5S_2 + 4.84G_5S_3 + 6.03G_5S_4 + 5.76G_5S_5 + 5.83G_5S_6 + 6.42G_5S_7 + 5.44G_5S_8 + 6.13G_5S_9 + 5.17G_5S_{10} + 5.48G_5S_{11} + 5.17G_5S_{12} + 5.61G_5S_{13} + 4.94G_5S_{14} + 4.06G_5S_{15} $$

#### *<b><u>Constraints:-</b></u>*

*Formulating the constraints basing on all the decision variables.*

##### *<u>Group Constraints:-</u>*

*Group1_Students:* $$  G_1S_1 +  G_1S_2 + G_1S_3 + G_1S_4 + G_1S_5 + G_1S_6 + G_1S_7 + G_1S_8 + G_1S_9 + G_1S_{10} + G_1S_{11} + G_1S_{12} + G_1S_{13} + G_1S_{14} + G_1S_{15} = 3 $$
*Group2_Students:* $$ G_2S_1 + G_2S_2 + G_2S_3 + G_2S_4 + G_2S_5 + G_2S_6 + G_2S_7 + G_2S_8 + G_2S_9 + G_2S_{10} + G_2S_{11} + G_2S_{12} + G_2S_{13} + G_2S_{14} + G_2S_{15} = 3 $$
*Group3_Students:* $$ G_3S_1 + G_3S_2 + G_3S_3 + G_3S_4 + G_3S_5 + G_3S_6 + G_3S_7 + G_3S_8 + G_3S_9 + G_3S_{10} + G_3S_{11} + G_3S_{12} + G_3S_{13} + G_3S_{14} + G_3S_{15} = 3 $$
*Group4_Students:* $$ G_4S_1 + G_4S_2 + G_4S_3 + G_4S_4 + G_4S_5 + G_4S_6 + G_4S_7 + G_4S_8 + G_4S_9 + G_4S_{10} + G_4S_{11} + G_4S_{12} + G_4S_{13} + G_4S_{14} + G_4S_{15} = 3 $$
*Group5_Students:* $$ G_5S_1 + G_5S_2 + G_5S_3 + G_5S_4 + G_5S_5 + G_5S_6 + G_5S_7 + G_5S_8 + G_5S_9 + G_5S_{10} + G_5S_{11} + G_5S_{12} + G_5S_{13} + G_5S_{14} + G_5S_{15} = 3 $$

##### *<u>Student:-</u>*

*Student1:* $$ G_1S_1 + G_2S_1 + G_3S_1 + G_4S_1 + G_5S_1 = 1 $$
*Student2:* $$ G_1S_2 + G_2S_2 + G_3S_2 + G_4S_2 + G_5S_2 = 1 $$
*Student3:* $$ G_1S_3 + G_2S_3 + G_3S_3 + G_4S_3 + G_5S_3 = 1 $$
*Student4:* $$ G_1S_4 + G_2S_4 + G_3S_4 + G_4S_4 + G_5S_4 = 1 $$
*Student5:* $$ G_1S_5 + G_2S_5 + G_3S_5 + G_4S_5 + G_5S_5 = 1 $$
*Student6:* $$ G_1S_6 + G_2S_6 + G_3S_6 + G_4S_6 + G_5S_6 = 1 $$
*Student7:* $$ G_1S_7 + G_2S_7 + G_3S_7 + G_4S_7 + G_5S_7 = 1 $$
*Student8:* $$ G_1S_8 + G_2S_8 + G_3S_8 + G_4S_8 + G_5S_8 = 1 $$
*Student9:* $$ G_1S_9 + G_2S_9 + G_3S_9 + G_4S_9 + G_5S_9 = 1 $$
*Student10:* $$ G_1S_{10} + G_2S_{10} + G_3S_{10} + G_4S_{10} + G_5S_{10} = 1 $$
*Student11:* $$ G_1S_{11} + G_2S_{11} + G_3S_{11} + G_4S_{11} + G_5S_{11} = 1 $$
*Student12:* $$ G_1S_{12} + G_2S_{12} + G_3S_{12} + G_4S_{12} + G_5S_{12} = 1 $$
*Student13:* $$ G_1S_{13} + G_2S_{13} + G_3S_{13} + G_4S_{13} + G_5S_{13} = 1 $$
*Student14:* $$ G_1S_{14} + G_2S_{14} + G_3S_{14} + G_4S_{14} + G_5S_{14} = 1 $$
*Student15:* $$ G_1S_{15} + G_2S_{15} + G_3S_{15} + G_4S_{15} + G_5S_{15} = 1 $$

##### *<u>Factor constraints:-</u>* 

*<u>For Factor GPA:</u>*

*Group1_Factor1:* $$ 2.93G_1S_1 +  3.68G_1S_2 + 3.11G_1S_3 + 3.82G_1S_4 + 3.91G_1S_5 + 2.57G_1S_6 + 3.29G_1S_7 + 3.84G_1S_8 + 3.33G_1S_9 + 3.18G_1S_{10} + 3.94G_1S_{11} + 3.18G_1S_{12} + 3.52G_1S_{13} + 3.36G_1S_{14} + 2.65G_1S_{15} \ge 3.35 $$
*Group2_Factor1:* $$ 2.93G_2S_1 + 3.68G_2S_2 + 3.11G_2S_3 + 3.82G_2S_4 + 3.91G_2S_5 + 2.57G_2S_6 + 3.29G_2S_7 + 3.84G_2S_8 + 3.33G_2S_9 + 3.18G_2S_{10} + 3.94G_2S_{11} + 3.18G_2S_{12} + 3.52G_2S_{13} + 3.36G_2S_{14} + 2.65G_2S_{15} \ge 3.35 $$
*Group3_Factor1:* $$ 2.93G_3S_1 + 3.68G_3S_2 + 3.11G_3S_3 + 3.82G_3S_4 + 3.91G_3S_5 + 2.57G_3S_6 + 3.29G_3S_7 + 3.84G_3S_8 + 3.33G_3S_9 + 3.18G_3S_{10} + 3.94G_3S_{11} + 3.18G_3S_{12} + 3.52G_3S_{13} + 3.36G_3S_{14} + 2.65G_3S_{15} \ge 3.35 $$
*Group4_Factor1:* $$ 2.93G_4S_1 + 3.68G_4S_2 + 3.11G_4S_3 + 3.82G_4S_4 + 3.91G_4S_5 + 2.57G_4S_6 + 3.29G_4S_7 + 3.84G_4S_8 + 3.33G_4S_9 + 3.18G_4S_{10} + 3.94G_4S_{11} + 3.18G_4S_{12} + 3.52G_4S_{13} + 3.36G_4S_{14} + 2.65G_4S_{15} \ge 3.35 $$
*Group5_Factor1:* $$ 2.93G_5S_1 + 3.68G_5S_2 + 3.11G_5S_3 + 3.82G_5S_4 + 3.91G_5S_5 + 2.57G_5S_6 + 3.29G_5S_7 + 3.84G_5S_8 + 3.33G_5S_9 + 3.18G_5S_{10} + 3.94G_5S_{11} + 3.18G_5S_{12} + 3.52G_5S_{13} + 3.36G_5S_{14} + 2.65G_5S_{15} \ge 3.35 $$

*<u>For Factor Technical Skill:</u>*

*Group1_Factor2:* $$ 9G_1S_1 +  5G_1S_2 + 4G_1S_3 + 6G_1S_4 + 10G_1S_5 + 9G_1S_6 + 8G_1S_7 + 8G_1S_8 + 10G_1S_9 + 8G_1S_{10} + 8G_1S_{11} + 7G_1S_{12} + 8G_1S_{13} + 6G_1S_{14} + 5G_1S_{15} \ge 7 $$ 
*Group2_Factor2:* $$ 9G_2S_1 + 5G_2S_2 + 4G_2S_3 + 6G_2S_4 + 10G_2S_5 + 9G_2S_6 + 8G_2S_7 + 8G_2S_8 + 10G_2S_9 + 8G_2S_{10} + 8G_2S_{11} + 7G_2S_{12} + 8G_2S_{13} + 6G_2S_{14} + 5G_2S_{15} \ge 7 $$
*Group3_Factor2:*$$ 9G_3S_1 + 5G_3S_2 + 4G_3S_3 + 6G_3S_4 + 10G_3S_5 + 9G_3S_6 + 8G_3S_7 + 8G_3S_8 + 10G_3S_9 + 8G_3S_{10} + 8G_3S_{11} + 7G_3S_{12} + 8G_3S_{13} + 6G_3S_{14} + 5G_3S_{15} \ge 7 $$
*Group4_Factor2:* $$ 9G_4S_1 + 5G_4S_2 + 4G_4S_3 + 6G_4S_4 + 10G_4S_5 + 9G_4S_6 + 8G_4S_7 + 8G_4S_8 + 10G_4S_9 + 8G_4S_{10} + 8G_4S_{11} + 7G_4S_{12} + 8G_4S_{13} + 6G_4S_{14} + 5G_4S_{15} \ge 7 $$
*Group5_Factor2:* $$ 9G_5S_1 + 5G_5S_2 + 4G_5S_3 + 6G_5S_4 + 10G_5S_5 + 9G_5S_6 + 8G_5S_7 + 8G_5S_8 + 10G_5S_9 + 8G_5S_{10} + 8G_5S_{11} + 7G_5S_{12} + 8G_5S_{13} + 6G_5S_{14} + 5G_5S_{15} \ge 7 $$

*<u>For Factor Teamwork:</u>*

*Group1_Factor3:* $$ 10G_1S_1 +  9G_1S_2 + 8G_1S_3 + 9G_1S_4 + 4G_1S_5 + 7G_1S_6 + 9G_1S_7 + 5G_1S_8 + 6G_1S_9 + 5G_1S_{10} + 5G_1S_{11} + 6G_1S_{12} + 6G_1S_{13} + 6G_1S_{14} + 5G_1S_{15} \ge 7 $$ 
*Group2_Factor3:* $$ 10G_2S_1 + 9G_2S_2 + 8G_2S_3 + 9G_2S_4 + 4G_2S_5 + 7G_2S_6 + 9G_2S_7 + 5G_2S_8 + 6G_2S_9 + 5G_2S_{10} + 5G_2S_{11} + 6G_2S_{12} + 6G_2S_{13} + 6G_2S_{14} + 5G_2S_{15} \ge 7 $$
*Group3_Factor3:*$$ 10G_3S_1 + 9G_3S_2 + 8G_3S_3 + 9G_3S_4 + 4G_3S_5 + 7G_3S_6 + 9G_3S_7 + 5G_3S_8 + 6G_3S_9 + 5G_3S_{10} + 5G_3S_{11} + 6G_3S_{12} + 6G_3S_{13} + 6G_3S_{14} + 5G_3S_{15} \ge 7 $$
*Group4_Factor3:* $$ 10G_4S_1 + 9G_4S_2 + 8G_4S_3 + 9G_4S_4 + 4G_4S_5 + 7G_4S_6 + 9G_4S_7 + 5G_4S_8 + 6G_4S_9 + 5G_4S_{10} + 5G_4S_{11} + 6G_4S_{12} + 6G_4S_{13} + 6G_4S_{14} + 5G_4S_{15} \ge 7 $$
*Group5_Factor3:* $$ 10G_5S_1 + 9G_5S_2 + 8G_5S_3 + 9G_5S_4 + 4G_5S_5 + 7G_5S_6 + 9G_5S_7 + 5G_5S_8 + 6G_5S_9 + 5G_5S_{10} + 5G_5S_{11} + 6G_5S_{12} + 6G_5S_{13} + 6G_5S_{14} + 5G_5S_{15} \ge 7 $$


#### *<b><u>Non-negativity of the decision variables:-</b></u>*

$$  G_iS_j \ge 0 \hspace{0.2cm} where \hspace{0.2cm} i=1,2,3,4,5 \hspace{0.2cm} and \hspace{0.2cm}j=1,2,...15 $$

#### *<b>2.Solve the Model in R: Implement the formulated model in R to find the optimal solution.</b>*

##### *Loading the required libraries.*
```{r}
library(lpSolveAPI)
```

##### *Creating an Linear Programming problem instance with 35 constraints and 75 decision variables.*
```{r}
lp <-make.lp(35, 75)
```

##### *Setting the objective function coefficients for the 75 decision variables.*
```{r}
set.objfn(lp, c(6.87,5.67,4.84, 6.03, 5.76, 5.83, 6.42, 5.44, 6.13, 5.17, 5.48, 5.17, 5.61, 4.94, 4.06, 
                6.87,5.67,4.84, 6.03, 5.76, 5.83, 6.42, 5.44, 6.13, 5.17, 5.48, 5.17, 5.61, 4.94, 4.06,  
                6.87,5.67,4.84, 6.03, 5.76, 5.83, 6.42, 5.44, 6.13, 5.17, 5.48, 5.17, 5.61, 4.94, 4.06,  
                6.87,5.67,4.84, 6.03, 5.76, 5.83, 6.42, 5.44, 6.13, 5.17, 5.48, 5.17, 5.61, 4.94, 4.06,  
                6.87,5.67,4.84, 6.03, 5.76, 5.83, 6.42, 5.44, 6.13, 5.17, 5.48, 5.17, 5.61, 4.94, 4.06))
```

##### *Setting the Objective sense to ‘max’.*
```{r}
lp.control(lp, sense = "max")
```

##### *Adding Group Constraints to the LP Problem.*
```{r}
set.row(lp,1, c(1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0))
set.row(lp,2, c(0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0))
set.row(lp,3, c(0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0))
set.row(lp,4, c(0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0))
set.row(lp,5, c(0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1))
```

##### *Adding Student Constraints to the LP Problem.*
```{r}
set.row(lp,6, c( 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0))
set.row(lp,7, c( 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0))
set.row(lp,8, c( 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0))
set.row(lp,9, c( 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0))
set.row(lp,10, c( 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0))
set.row(lp,11, c( 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0))
set.row(lp,12, c( 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0))
set.row(lp,13, c( 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0))
set.row(lp,14, c( 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0))
set.row(lp,15, c(0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0))
set.row(lp,16, c(0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0))
set.row(lp,17, c(0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0))
set.row(lp,18, c( 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0))
set.row(lp,19, c( 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0))
set.row(lp,20, c( 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1))
```

##### *Adding Factor constraints to the LP Problem*
```{r}
#Factor_1 Constraints:-

set.row(lp,21, c(2.93, 3.68, 3.11, 3.82, 3.91, 2.57, 3.29, 3.84, 3.33, 3.18, 3.94, 3.18, 3.52, 3.36, 2.65, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0))
set.row(lp,22, c( 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 2.93, 3.68, 3.11, 3.82, 3.91, 2.57, 3.29, 3.84, 3.33, 3.18, 3.94, 3.18, 3.52, 3.36, 2.65, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0))
set.row(lp,23, c(0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 2.93, 3.68, 3.11, 3.82, 3.91, 2.57, 3.29, 3.84, 3.33, 3.18, 3.94, 3.18, 3.52, 3.36, 2.65, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0))
set.row(lp,24, c( 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 2.93, 3.68, 3.11, 3.82, 3.91, 2.57, 3.29, 3.84, 3.33, 3.18, 3.94, 3.18, 3.52, 3.36, 2.65, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0))
set.row(lp,25, c( 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 2.93, 3.68, 3.11, 3.82, 3.91, 2.57, 3.29, 3.84, 3.33, 3.18, 3.94, 3.18, 3.52, 3.36, 2.65))

#Factor_2 Constraints:-

set.row(lp,26, c(9, 5, 4, 6, 10, 9, 8, 8, 10, 8, 8, 7, 8, 6, 5, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0))
set.row(lp,27, c( 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 9, 5, 4, 6, 10, 9, 8, 8, 10, 8, 8, 7, 8, 6, 5, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0))
set.row(lp,28, c(0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 9, 5, 4, 6, 10, 9, 8, 8, 10, 8, 8, 7, 8, 6, 5, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0))
set.row(lp,29, c( 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 9, 5, 4, 6, 10, 9, 8, 8, 10, 8, 8, 7, 8, 6, 5, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0))
set.row(lp,30, c( 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 9, 5, 4, 6, 10, 9, 8, 8, 10, 8, 8, 7, 8, 6, 5))

#Factor_3 Constraints:-

set.row(lp,31, c(10, 9, 8, 9, 4, 7, 9, 5, 6, 5, 5, 6, 6, 6, 5, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0))
set.row(lp,32, c( 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 10, 9, 8, 9, 4, 7, 9, 5, 6, 5, 5, 6, 6, 6, 5, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0))
set.row(lp,33, c(0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 10, 9, 8, 9, 4, 7, 9, 5, 6, 5, 5, 6, 6, 6, 5, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0))
set.row(lp,34, c( 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 10, 9, 8, 9, 4, 7, 9, 5, 6, 5, 5, 6, 6, 6, 5, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0))
set.row(lp,35, c( 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 10, 9, 8, 9, 4, 7, 9, 5, 6, 5, 5, 6, 6, 6, 5))
```

##### *Creating a vector rhs containing the right-hand side values for the two constraints.*
```{r}
rhs<-c(3,3,3,3,3,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,3.35,3.35,3.35,3.35,3.35,7,7,7,7,7,7,7,7,7,7)
```

##### *Setting the right-hand side values for the constraints in the LP problem.*
```{r}
set.rhs(lp,rhs)
```

##### *Setting the constraint types for the constraints.*
```{r}
set.constr.type(lp,c("=","=","=","=","=","=","=","=","=","=","=","=","=","=","=","=","=","=","=", "=",">=",">=",">=", ">=",">=",">=",">=",">=",">=",">=",">=",">=",">=",">=",">="))
```

##### *Setting bounds for the decision variables explicitly.*
```{r}
set.bounds(lp,lower = rep(0,75))
```

##### *Assigning names to the decision variables.*
```{r}
lp.colnames <- c("G1S1", "G1S2", "G1S3", "G1S4", "G1S5", "G1S6", "G1S7", "G1S8", "G1S9", "G1S10", "G1S11", "G1S12", "G1S13", "G1S14", "G1S15", "G2S1", "G2S2", "G2S3", "G2S4", "G2S5", "G2S6", "G2S7", "G2S8", "G2S9", "G2S10", "G2S11", "G2S12", "G2S13", "G2S14", "G2S15", "G3S1", "G3S2", "G3S3", "G3S4", "G3S5", "G3S6", "G3S7", "G3S8", "G3S9", "G3S10", "G3S11", "G3S12", "G3S13", "G3S14", "G3S15", "G4S1", "G4S2", "G4S3", "G4S4", "G4S5", "G4S6", "G4S7", "G4S8", "G4S9", "G4S10", "G4S11", "G4S12", "G4S13", "G4S14", "G4S15", "G5S1", "G5S2", "G5S3", "G5S4", "G5S5", "G5S6", "G5S7", "G5S8", "G5S9", "G5S10", "G5S11", "G5S12", "G5S13", "G5S14", "G5S15")
```

##### *Assigning names to the constraints.*
```{r}
lp.rownames <- c("Group1Students", "Group2Students", "Group3Students", "Group4Students", "Group5Students", "Student1", "Student2", "Student3", "Student4", "Student5", "Student6", "Student7", "Student8", "Student9", "Student10","Student11","Student12", "Student13", "Student14", "Student15", "Group1Factor1", "Group2Factor1", "Group3Factor1","Group4Factor1", "Group5Factor1", "Group1Factor2","Group2Factor2","Group3Factor2","Group4Factor2", "Group5Factor2", "Group1Factor3","Group2Factor3","Group3Factor3","Group4Factor3", "Group5Factor3")
dimnames(lp) = list(lp.rownames,(1:75))
```

##### *Setting type for the variables to 'Integer'.*
```{r}
set.type(lp,1:75,type = "integer")
lp
```

##### *Solving the LP problem using the specified constraints and objective function.*
```{r}
solve(lp)
```

##### *Retrieving the optimal objective value of the LP problem.*
```{r}
get.objective(lp)
```

##### *Retrieving the values of the decision variables at the optimal solution.*

```{r}
get.variables(lp)
```

##### *Retrieving the values of the constraints at the optimal solution.*
```{r}
get.constraints(lp)
```


